This report documents a project regarding the
Guarded Command Language (GCL), presented in the course 02242
Program Analysis, DTU, in the fall of 2010.
The project goal is to design, implement and test
a program analysis for GCL, including a parser,
worklist algorithm, several analyses and transformations.
As part of the project, a comparison between the Succint Solver
presented in the course and the implemented worklist algorithm
must be made and documented.
The project has been implemented in Java.

\paragraph{Report overview}

The report includes an introduction to the report, following by the main
content of the report. The content of the project has been divided into
3 main parts: Analysis, Design and Implementation, and Test.

Analysis describes the theoretical background
of the project. This includes the general definitions in
the project, as well as the 3 primary analyses used in
the project: Reaching Definitions, Live Variables and
Constant Propagation.

Design and Implementation introduces the design
and implementation of several parts of the program
analysis module. This includes the data structure used to
represent the GCL programs internally, the worklist algorithm
used as part of a framework for calculating the analyses,
as well as the three main transformations based on
the analyses described earlier: Program Slicing,
Dead Code Elimination and Constant Folding.

Test describes the testing methods used in
the project, the test results, issues regarding testing,
and the conclusion of the tests in regards to the
main transformations. Finally, the test comparison
with the Succinct solver is described here.

The report then ends with an evaluation, discussing the
results and conclusions, and a conclusion, describing
what has been attained and future work and considerations.

